### 2024-06-05 Scripts/Manuscript_scripts.Rmd
Update Manuscript_scripts.Rmd

>>>>

```{r Fig 3c}

```{r Fig 3d}

df_vf_num = read.csv(r"(LRT_microbiomes_critical_illness\Data_repository\Main_figures\vfdb_profile-species_all-vf_num.csv)")
df_vf_num$species <- factor(df_vf_num$species,levels = c('pseudomonas','acinetobacter','klebsiella','corynebacterium','stenotrophomonas'))
my_comparisons <- list(c('pseudomonas','stenotrophomonas'),c('pseudomonas','corynebacterium'),
                       c('acinetobacter','stenotrophomonas'),c('acinetobacter','corynebacterium'),
                       c('klebsiella','stenotrophomonas'),c('klebsiella','corynebacterium'))

map <-c('acinetobacter'=3.965,'pseudomonas'=6.644,'klebsiella'=5.596,'corynebacterium'=2.841,'stenotrophomonas'=4.613)
df_vf_num <- df_vf_num %>%
  mutate(genome_length = map[species])
df_vf_num <- df_vf_num %>%
  mutate(vf_num_normalized = vf_num/genome_length)

ggboxplot(df_vf_num, x="species", y="vf_num_normalized",
          color = "species", palette =c("#E41A1C","#377EB8","#4DAF4A","#984EA3","#FF7F00","#F781BF","#A65628","#F781BF"),
          add = "jitter",add.params = list(size = 1.5, alpha = 0.5))  +
  stat_boxplot(mapping = aes(color=species),geom = "errorbar",width=0.1,lwd=0.8) +
  stat_compare_means(comparisons = rev(my_comparisons),label = "p.label",size=10,
                     vjust=.15,tip.length= c(0,0,0)) +
  scale_x_discrete(labels=c('P. aeruginosa','A. baumannii','K. pneumoniae','C. striatum','S. maltophilia'))+
  labs(x=NULL,y='Number of VFs (Normalized)',) +  
  theme_pubr(base_size = 32,legend ='none',x.text.angle = 30)

```{r Fig 3e}

```{r Fig 3f}

```{r Fig S3a}

df_argsubtype_top3 <- read.csv(r"(LRT_microbiomes_critical_illness\Data_repository\Supplementary_figures\statistics-pneumonia_patients-argsubtypes_top-each_hospital.csv)")
df_argsubtype_top3$arg_type <- factor(df_argsubtype_top3$arg_type,levels = c('Multidrug','Beta-lactam','Aminoglycoside','MLS'))

pal <- c('Multidrug__acra'='#3182BD','Multidrug__multidrug_transporter'='#00A2CE','Beta-lactam__shv-39'='#31A354','Beta-lactam__cfxa2'='#008179',
         'Beta-lactam__oxa-225'='#9ED1A6','Aminoglycoside__rmtb'='#FFB500',"Aminoglycoside__aac(6')-i"='#FFB686',"Aminoglycoside__aph(3'')-i"='#FFEDCB',
         'MLS__maca'='#EFD7EB','MLS__ermx'='#FFD1E0')
ggbarplot(df_argsubtype_top3,
          x='sites',y='mean_abundance',fill = 'argsubtypes') + 
  scale_fill_manual(values = pal)  +
  labs(x=NULL,y='Mean abundance') +
  guides(fill=guide_legend('ARG subtype',ncol = 2)) +
  facet_grid(~arg_type) + theme_pubr(border = T,x.text.angle = 30,legend = 'bottom',base_size = 20)

```{r Fig S3b}

df_argtypes_collection <- read.csv(r"(LRT_microbiomes_critical_illness\Data_repository\Supplementary_figures\argtypes-collection_all_sites.csv)")
df_argtypes_collection2 <- arrange(df_argtypes_collection,SamplingSite)
my_comparisons <- list( c("HospitalA", "HospitalB"), c("HospitalA", "HospitalC"), c("HospitalB", "HospitalC") )

#plot for 'MLS'
ggboxplot(df_argtypes_collection2, x="SamplingSite", y="macrolide.lincosamide.streptogramin",
          color = "SamplingSite", palette =c("#f93800", "#007a8b","#ffb500"),
          add = "jitter",add.params = list(size = 1.5, alpha = 0.5))  +
  stat_boxplot(mapping = aes(color=SamplingSite),geom = "errorbar",width=0.1)  +
  stat_compare_means(comparisons = my_comparisons,label = "p.label",size=5,
                     vjust=.15,tip.length= c(0,0,0)) + 
  yscale("log10", .format = F)  + 
  labs(x=NULL,y='ARG abundance',) +  ggtitle('MLS') +
  theme_pubr(base_size = 18,legend ='none',x.text.angle = 30) +
  ggeasy::easy_center_title()

#plot for 'tetracycline'
ggboxplot(df_argtypes_collection2, x="SamplingSite", y="tetracycline",
          color = "SamplingSite", palette =c("#f93800", "#007a8b","#ffb500"),
          add = "jitter",add.params = list(size = 1.5, alpha = 0.5))  +
  stat_boxplot(mapping = aes(color=SamplingSite),geom = "errorbar",width=0.1)  +
  stat_compare_means(comparisons = my_comparisons,label = "p.label",size=5,
                     vjust=.15,tip.length= c(0,0,0)) + 
  yscale("log10", .format = F)  + 
  labs(x=NULL,y='ARG abundance',) +  ggtitle('Tetracycline') +
  theme_pubr(base_size = 18,legend ='none',x.text.angle = 30) +
  ggeasy::easy_center_title()

#plot for 'sulfonamide'
ggboxplot(df_argtypes_collection2, x="SamplingSite", y="sulfonamide",
          color = "SamplingSite", palette =c("#f93800", "#007a8b","#ffb500"),
          add = "jitter",add.params = list(size = 1.5, alpha = 0.5))  +
  stat_boxplot(mapping = aes(color=SamplingSite),geom = "errorbar",width=0.1)  +
  stat_compare_means(comparisons = my_comparisons,label = "p.label",size=5,
                     vjust=.15,tip.length= c(0,0,0)) + 
  yscale("log10", .format = F)  + 
  labs(x=NULL,y='ARG abundance',) +  ggtitle('Sulfonamide') +
  theme_pubr(base_size = 18,legend ='none',x.text.angle = 30) +
  ggeasy::easy_center_title()

```{r Fig S3c}

df_nmds = read.csv(r'(LRT_microbiomes_critical_illness\Data_repository\Supplementary_figures\NMDS_arg_subtypes-all_sites.csv)',header = TRUE)
my_comparisons <- list( c("HospitalA", "HospitalB"), c("HospitalA", "HospitalC"), c("HospitalB", "HospitalC") )
df_nmds$NMDS1 <- as.numeric(df_nmds$NMDS1)
df_nmds$NMDS2 <- as.numeric(df_nmds$NMDS2)

#scatter plot to show the argsubtypes distribution for all ETA samples
p_scatter <- ggscatter(df_nmds,
                       x = "NMDS1", 
                       y = "NMDS2",
                       color = 'SamplingSite',
                       palette = c("#e64833", "#01a188", "#3e5386"),
                       #shape = "SamplingSite",
                       size=5,
                       ggtheme = theme_pubr(border = T,base_size = 32),
                       legend = 'none')  + xlim2(limits = c(-2.5,2.5)) + ylim2(limits = c(-2.5,1.8)) +  
  theme(axis.title = element_blank(),
        axis.ticks = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank())
p <-ggscatterhist(df_nmds,
              x = "NMDS1", 
              y = "NMDS2",
              xlim= c(-2.2,2.2),
              ylim= c(-2.5,1.5),
              color = "SamplingSite",
              size=5,
              palette = c("#e64833", "#01a188", "#3e5386"),
              margin.plot = "density",
              ggtheme = theme_pubr(border = T,base_size = 32),
              margin.params = list(fill="SamplingSite",size=0.3),
              legend = 'right')

```{r Fig S3d}

df_splsda = read.csv(r'(LRT_microbiomes_critical_illness\Data_repository\Supplementary_figures\PLS-DA_arg_subtypes-all_sites.csv)',header = TRUE)

### Plot the PLS-DA components distribution 
plot_splsda <- ggscatter(df_splsda,
          x = "comp1", 
          y = "comp2",
          color = "SamplingSite",
          # shape = "SamplingSite",
          size=4,
          ellipse=TRUE,
          # star.plot = TRUE,
          # palette = c("#00AFBB", "#E7B800", "#FC4E07"), #palette1
          palette = c("#c23726", "#00C19B", "#2c83c6"),  #palette2
          # palette = c("#e64833", "#01a188", "#3e5386"), #palette3
          # margin.plot = "boxplot",
          # ggtheme = theme_pubr(),
          # margin.params = list(fill="SamplingSite",size=0.36),
          legend = 'right')
ggpar(plot_splsda,
      xlab = "Component 1 (12%)",
      ylab = "Component 2 (3%)",
      font.xtickslab=c(32),
      font.ytickslab=c(32),
      font.x = c(32),
      font.y = c(32),
      legend.title = "Sampling site",
      font.legend = c(32))  + theme_pubr(base_size = 32,legend = "none",margin = FALSE,border = T)

```{r Fig S3e}

df_loading1 <- read.csv(r'(LRT_microbiomes_critical_illness\Data_repository\Supplementary_figures\PLS-DA_arg_subtypes-all_sites-loadings-comp1.csv)',header = TRUE)
plot_loading1 <- ggbarplot(df_loading1[1:6,],x = "X", y = "value.var",orientation = "horiz",
          fill = "#e6a024", color = "#e6a024",sort.val = 'desc',ggtheme = theme_bw(base_size = 19)) + labs(title = 'Component-1') +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text = element_blank(),
        axis.title = element_blank())

plot_loading1 <- ggpar(plot_loading1,
      orientation = 'reverse')  #+ theme_pubr(base_size = 32,legend = "right",margin = FALSE)

#plot loadings on component-2
df_loading2 <- read.csv(r'(LRT_microbiomes_critical_illness\Data_repository\Supplementary_figures\PLS-DA_arg_subtypes-all_sites-loadings-comp2.csv)',header = TRUE)
df_loading2 <- df_loading2 %>%
  mutate(X=str_replace(X,'_chloramphenicol acetyltransferase',''),
         X=str_replace(X,'_\\(ramA_or_sud2\\)',''))

plot_loading2 <- ggbarplot(df_loading2[1:6,],x = "X", y = "value.var",orientation = "horiz",
                           fill = "#5bb4e5", color = "#5bb4e5",sort.val = 'asc',ggtheme = theme_bw(base_size = 19))+ labs(title = 'Component-2') +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text = element_blank(),
        axis.title = element_blank())

plot_loading1 | plot_loading2

```{r Fig S3f}

lmeRes <- read.csv(r'(LRT_microbiomes_critical_illness\Data_repository\Supplementary_figures\lme-argsubtypes.csv)')

# Subset and reformat the dataset
lmeRes <- subset(lmeRes, predictor %in% c("Sampling_Day_ICU", "Oxygenation.Index.of.The.Day", "Diagnosed.PneumoniaYes"))
lmeRes <- lmeRes %>%
  mutate(predictor_reformat = recode(predictor, 
                            "Sampling_Day_ICU" = "Days in the ICU", 
                            "Oxygenation.Index.of.The.Day" = "Oxygenation index", 
                            "Diagnosed.PneumoniaYes" = "Diagnosed pneumonia"))

# Setting the factor levels
lmeRes$predictor_reformat <- factor(lmeRes$predictor_reformat, levels = c("Diagnosed pneumonia", "Days in the ICU", "Oxygenation index"))

valid_x <- c("Days in the ICU", "Oxygenation index", "Diagnosed pneumonia")

multiVarSigDF <- subset(lmeRes, predictor_reformat %in% valid_x & p_adjusted <= 0.05)
multiVarSigDF$p_adjusted_log <- -log10(multiVarSigDF$p_adjusted)

uniqueVar <- multiVarSigDF[, c('item', 'predictor_reformat', 'p_adjusted_log', 'p_adjusted')]
spW <- uniqueVar %>% group_by(item) %>% summarise(value = sum(p_adjusted_log))
xW <- uniqueVar %>% group_by(predictor_reformat) %>% summarise(value = sum(p_adjusted_log))

YsName <- spW$item[order(spW$value)]
# XsName <- xW$predictor_reformat[order(match(xW$predictor_reformat, valid_x))]
XsName <- rev(c("Diagnosed pneumonia", "Days in the ICU", "Oxygenation index"))

ord <- c(XsName, YsName)

grid.col <- c(rep('#FDCA4E' %>% alpha(., 1), length(YsName)),
              rep('#89ADD4' %>% alpha(., 1), length(XsName)))
names(grid.col) <- c(YsName, XsName)

color_lookup <- c(
  "Diagnosed pneumonia" = alpha("#EB9E9A", 0.7),
  "Days in the ICU" = alpha("#BAA8CE", 0.7),
  "Oxygenation index" = alpha("#DFE7F1", 0.7)
)

col_fun <- data.frame(
  uniqueVar$item,
  uniqueVar$predictor_reformat
)
col_fun$col <- color_lookup[uniqueVar$predictor_reformat]
col_fun$col[is.na(col_fun$col)] <- '#5D8DC3' %>% alpha(., 0.5)  # Default color for predictors not in the lookup

```{r Fig S3g}

```{r Fig S3h}

```{r Fig S3f}
